---
title: "p8130_hw2_brm2150"
author: "Brooklynn McNeil"
date: "2024-09-29"
output: html_document
---

This document if for the Homework Assignment #2 for Biostatistical Methods 1.

## Problem 1

The probability of having at least 1 dental checkup in a two-year period is 73%. 56 random individuals are participating in a study.

a.  What is the probability of exactly 40 people having at least 1 dental checkup?

Use the probability mass function (pmf) for a binomial distribution.

n = 56 (number of trials)

k = 40 (number of successes)

p = 0.73 (probability of success)

P(X = k) = (n k)p^k^ (1-p)^n-k^

(n k) = n! / (k!(n-k)!) = 56!/(40!(56-40)! = 56!/(40! x 16!)

P(X = k) = 56!/(40! x 16!) x 0.73^40^ x (1-0.73)^(56-40)^

P(X = k) = **0.1133 or 11.33%** chance that exactly 40 people will get at least 1 dental checkup in the next 2 years.

b.  what is the probability of at least 40 people having at least 1 dental checkup.

```{r}
# P(X>=40) = 1 - P(X<40)

n = 56
p = 0.73
k = 39

X = (1 - dbinom(k, n, p))

X
```

c.  Since the number of trials is large and the probability of success is high, we can use a normal distribution to approximate the probabilities from above.

```{r}
# normal approximation for part a
n = 56
p = 0.73  
mu <- n * p
sigma = sqrt(n * p * (1 - p))

exact_40 = pnorm(40, mean = mu, sd = sigma)

exact_40

# normal approximation for part b

at_least_40 = 1 - pnorm(39, mean = mu, sd = sigma)

at_least_40

```

d\. How many participants are expected have at least 1 checkup?

```{r}
# use the expected value formula

n = 56
p = 0.73

exp_val = n*p

exp_val
```

e\. What is the standard deviation of people that will have at least 1 checkup.

```{r}
n = 56
p = 0.73

sd_x = sqrt(
  (n * p)*(1 - p)
)

sd_x
```

## Problem 2

The number of tornadoes in the United States follows a poisson distribution with lambda = 6 (tornadoes per year).

a\. What is the probability of having fewer than 3 tornadoes next year?

probability mass function for poisson distribution:

Y = lambda

P(X = k) = (Y^k^ e^-Y^ )/k!

```{r}

lambda = 6

torn0 = dpois(0, lambda = lambda)
torn1 = dpois(1, lambda = lambda)
torn2 = dpois(2, lambda = lambda)

torn0 + torn1 + torn2
```

b\. What is the probability of having exactly 3 tornadoes next year?

```{r}
k = 3
lambda = 6

torn3 = dpois(k, lambda = lambda)

torn3
```

c\. What is the probability of having more than 3 tornadoes next year?

```{r}
# P(X) = P(1-P(k))
# 1 - P(3 or less tornadoes ocurring)

1 - (torn0+torn1+torn2+torn3)
```

## Problem 3

Systolic blood pressure for men 20-29 years old in the US is normally distributed. The population mean in 128.0 and the standard deviation is 10.2

a\. The probability that a randomly selected American male 20-29 yrs has systolic blood pressure above 137.0 can be solved with the normal distribution probability function.

P(X \> 137.0)

```{r}
mu = 128.0
sigma = 10.2
q = 137.0

p_sysbp_137 = pnorm(q, mean = mu, sd = sigma, lower.tail = FALSE)

p_sysbp_137
```

b. The probability that the sample mean for the blood pressure of 50 males will be less than 125.0 is performed by first calculating the standard error and using that in the normal distribution probability function.

```{r}
mu = 128.0
sigma = 10.2
n = 50
q = 125.0 

std_err = sigma/sqrt(n)

p_sysbp_125 = pnorm(q, mean = mu, sd = std_err, lower.tail = TRUE)

p_sysbp_125
```

c\. The 90th percentile of the sampling distribution of size n = 40

```{r}
mu = 128.0
sigma = 10.2
n = 40
p = 0.90

std_err = sigma/sqrt(n)

pct_90 = qnorm(p, mean = mu, sd = std_err)
pct_90
```

## Problem 4
