---
title: "p8130_hw2_brm2150"
author: "Brooklynn McNeil"
date: "2024-09-29"
output: html_document
---

This document if for the Homework Assignment #2 for Biostatistical Methods 1.

## Problem 1

The probability of having at least 1 dental checkup in a two-year period is 73%. 56 random individuals are participating in a study.

a.  What is the probability of exactly 40 people having at least 1 dental checkup?

Use the probability mass function (pmf) for a binomial distribution.

n = 56 (number of trials)

k = 40 (number of successes)

p = 0.73 (probability of success)

P(X = k) = (n k)p^k^ (1-p)^n-k^

(n k) = n! / (k!(n-k)!) = 56!/(40!(56-40)! = 56!/(40! x 16!)

P(X = k) = 56!/(40! x 16!) x 0.73^40^ x (1-0.73)^(56-40)^

P(X = k) = **0.1133 or 11.33%** chance that exactly 40 people will get at least 1 dental checkup in the next 2 years.

b.  what is the probability of at least 40 people having at least 1 dental checkup.

```{r}
# P(X>=40) = 1 - P(X<40)

n = 56
p = 0.73
k = 39

X = (1 - dbinom(k, n, p))

X
```

c\. Since the number of trials is large and the probability of success is high, we can use a normal distribution to approximate the probabilities from above.

```{r}
# normal approximation for part a
n = 56
p = 0.73  
mu <- n * p
sigma = sqrt(n * p * (1 - p))

exact_40 = pnorm(40, mean = mu, sd = sigma)

exact_40

# normal approximation for part b

at_least_40 = 1 - pnorm(39, mean = mu, sd = sigma)

at_least_40

```

d\. How many participants are expected have at least 1 checkup?

```{r}
# use the expected value formula

n = 56
p = 0.73

exp_val = n*p

exp_val
```

e\. What is the standard deviation of people that will have at least 1 checkup.

```{r}
n = 56
p = 0.73

sd_x = sqrt(
  (n * p)*(1 - p)
)

sd_x
```

## Problem 2

## Problem 3

## Problem 4
